Data Specialization for a Staged Lambda Calculus

### 2015-02-04 paper/figures/splitting-overview.tex
major reorganization of section 2 (in progress) - moving lots of text around!
2-evaluation -> residualization
elaborated on the splitting overview figures

@@

\textbf{\lang\ Dynamic Semantics:}
\begin{itemize}
\item 
\bbone-Evaluation: $e\mathbin{\redonesym}[\xi;v]$, where $\xi$ is a \emph{residual table} and $v$ is a \emph{partial value}.

\item
\bbtwo-Evaluation: $e\mathbin{\redtwosym}q$, where $q$ is a \emph{residual} in the unstaged language \langTwo

@@

\textbf{\lang\ Splitting:}

\hspace{2em}\bbone-Splitting Structure: $e \splitonesym [c,l.r]$, where:

\hspace{4em}$c$ is the \emph{combined term}
(representing all stage~\bbone\ subcomputations in $e$)
in \langmono

\hspace{4em}$l.r$ is the \emph{resumer}
(representing all stage~\bbtwo\ subcomputations in $e$)
in \langmono

\hspace{4em}$c~\redsym~(y,b)$, where 
$y$ is the \emph{\bbone-result} of $e$, and
$b$ is the \emph{boundary value} of $e$

\hspace{2em}\bbone-Splitting Correctness: If $e\mathbin{\redonesym}[\xi;v]$, then:

\hspace{4em}$y$ is identical to $\masko{\xi;v}$

\hspace{4em}$\maskt{\xi;v}$ and \texttt{(let l=b in r)} 
reduce (via $\redsym$) to equivalent values.

\hspace{2em}\bbtwo-Splitting Structure: $e \splittwosym [p,l.r]$, where:

\hspace{4em}$p$ is the \emph{precomputation}
(representing all stage~\bbone\ subcomputations in $e$)
in \langmono

\hspace{4em}$l.r$ is the \emph{resumer}
(representing all stage~\bbtwo\ subcomputations in $e$)
in \langmono

\hspace{2em}\bbtwo-Splitting Correctness:
If $e~\redtwosym~q$, and $q~\redsym~v$, then
\texttt{(let l=p in r)}~$\redsym~v'$ and $v,v'$ are equivalent.
% TODO: FIX THIS
% if e reduces to q, then p reduces to u
% and q == [u/l]r
% that is, stage 1 reduces, even if stage 2 doesn't

\caption{Summary of \lang\ evaluation and splitting.}
\label{fig:termSplitSummary}
\end{figure*}

>>>>

%!TEX root = ../paper.tex

@@

\textbf{\lang\ Dynamic Semantics (\ref{sec:semantics})}
\begin{itemize}
\item 
\bbone-Evaluation: $e\mathbin{\redonesym}[\xi;v]$, where $\xi$ is a \emph{residual table} and $v$ is a \emph{partial value}.

\item
Residualization: $e\mathbin{\redtwosym}q$, where $q$ is a \emph{residual} in the unstaged language \langTwo

@@

\caption{Summary of \lang\ evaluation.}
\label{fig:termSplitSummary}
\end{figure*}

\begin{figure*}
\begin{abstrsyn}

\textbf{Single-Output Pipelines:}

\hspace{2em} Construction: $\pipeS p l r$, where

	\hspace{4em} $p$ is the {\em precomputation}, a monostage term representing the stage \bbone\ part of the program

	\hspace{4em} $l.r$ is the {\em resumer}, a monostage term representing the stage \bbtwo\ part of the program

\hspace{2em} Evaluation Form: $\sepredtwo {\pipeS p l r} q$, where

	\hspace{4em} $\pipeS p l r$ is the single-output pipeline to be evaluated

	\hspace{4em} $q$ is the resulting residual, a monostage term representing remaining stage \bbtwo\ work

\hspace{2em} Evaluation Rule: $\sepredtwo {\pipeS p l r} {[b/l]r}$, where $b$ (called the {\em boundary value}) is generated by $\reduce p b$

\textbf{Mutiple-Output Pipelines:}

\hspace{2em} Construction: $\pipeM c l r$, where

	\hspace{4em} $c$ is the {\em combined result}, a monostage term representing the stage \bbone\ part of the program

	\hspace{4em} $l.r$ is the {\em resumer}, a monostage term representing the stage \bbtwo\ part of the program

\hspace{2em} Evaluation Form: $\sepredone {\pipeM c l r} {\mval i q}$, where

	\hspace{4em} $\pipeM c l r$ is the multi-output pipeline to be evaluated

	\hspace{4em} $\mval i q$ is the resulting masked value (defined below)

\hspace{2em} Evaluation Rule: $\sepredtwo {\pipeM c l r} {\mval i {[b/l]r}}$, where $b$ (also the {\em boundary value}) is generated by $\reduce c (i,b)$

\textbf{Masked Values:}

\hspace{2em} Construction: $\mval i q$, where

	\hspace{4em} $i$ is the {\em immediate result}, a monostage value representing the stage \bbone\ result

	\hspace{4em} $q$ is the {\em residual}, a monostage term representing stage \bbtwo\ work left to do

\end{abstrsyn}
\caption{Overview of Output Language.}
\label{fig:outputTermSummary}
\end{figure*}

\setlength{\unitlength}{2.8pt}
\begin{figure*}
\begin{abstrsyn}
\centering
\begin{tabular}{c|c}

{\Large \bf Splitting at World \bbonem} & 
{\Large \bf Splitting at World \bbtwo} \\ \hline

\begin{minipage}[t]{0.45\textwidth}
Splitting Terms: $(\colmix e A) \splitonesym \pipeM p l r$, where
\par \hspace{2em} $\colmix e A$ is the \lang\ input term to be split 
\par \hspace{2em} $\pipeM p l r$ is the resulting multi-output pipeline
\\
Masking Partial Values: $[\xi;v] \vsplito \mval i q$, where
\par \hspace{2em} $[\xi;v]$ is a partial value in \lang, resulting from $\redonesym$
\par \hspace{2em} $\mval i q$ is the resulting masked value
\end{minipage}
&

\begin{minipage}[t]{0.45\textwidth}
Splitting Terms: $(\coltwo e A) \splittwosym \pipeS p l r$, where 	
\par \hspace{2em} $\coltwo e A$ is the \lang\ input term to be split
\par \hspace{2em} $\pipeS p l r$ is the resulting single-output pipeline
\end{minipage}

\\

\begin{picture} (80,54) (10,30)

\thicklines
\put(25,37){\oval(22,10)}
\put(25,75){\oval(22,10)}
\put(75,37){\oval(22,10)}
\put(75,75){\oval(22,10)}
\put(36,37){\vector(1,0){28}}
\put(36,75){\vector(1,0){28}}
\put(25,70){\vector(0,-1){28}}
\put(75,70){\vector(0,-1){28}}

\put(14,75){\raisebox{-0.5ex}{\makebox[22 \unitlength]{$\colone e A$}}}
\put(14,37){\raisebox{-0.5ex}{\makebox[22 \unitlength]{$[\xi,v]$}}}
\put(64,75){\raisebox{-0.5ex}{\makebox[22 \unitlength]{$\pipeM c l r$}}}
\put(64,37){\raisebox{-0.5ex}{\makebox[22 \unitlength]{$\mval i q$}}}

\put(25,78){\makebox[50 \unitlength]{Splitting $\left(\splitonesym\right)$}}
\put(25,32){\makebox[50 \unitlength]{Masking $\left(\vsplito\right)$}}
\put(76,55){\parbox[l]{20 \unitlength}{Separated-Form \\ Evaluation $\left(\sepredonesym\right)$}}
\put(9,55){\parbox[r]{20 \unitlength}{Interleaved \\ Evaluation $\left(\redonesym\right)$}}

\put(28,66){\parbox[t]{44 \unitlength}{ 
	\textbf{Correctness Condition:} \\ 
	Splitting followed by evaluation (top and right arrows)
	should produce the same result as 
	evaluation followed by masking (left and bottom arrows),
	up to a term equivalence.
}}
\end{picture}

&
\begin{picture} (80,54) (10,30)

\thicklines
\put(25,75){\oval(22,10)}
\put(75,37){\oval(22,10)}
\put(75,75){\oval(22,10)}
\put(25,37){\vector(1,0){39}}
\put(36,75){\vector(1,0){28}}
\put(25,70){\line(0,-1){33}}
\put(75,70){\vector(0,-1){28}}

\put(14,75){\raisebox{-0.5ex}{\makebox[22 \unitlength]{$\coltwo e A$}}}
\put(64,75){\raisebox{-0.5ex}{\makebox[22 \unitlength]{$\pipeS p l r$}}}
\put(64,37){\raisebox{-0.5ex}{\makebox[22 \unitlength]{$q$}}}

\put(25,78){\makebox[50 \unitlength]{Splitting $\left(\splittwosym\right)$}}
\put(76,55){\parbox[l]{20 \unitlength}{Separated-Form \\ Evaluation $\left(\sepredtwosym\right)$}}
\put(9,55){\parbox[r]{20 \unitlength}{Interleaved \\ Evaluation $\left(\redtwosym\right)$}}

\put(28,66){\parbox[t]{44 \unitlength}{ 
	\textbf{Correctness Condition:} \\ 
	Splitting followed by evaluation (top and right arrows)
	should produce the same result as 
	residualization (left and bottom arrow),
	up to a term equivalence.
}}
\end{picture}

\\

\end{tabular}

\end{abstrsyn}
\caption{Overview of Splitting.}
\label{fig:outputTermSummary}
\end{figure*}
